---
title: "Tidybiology +DS: Session 2 "
subtitle: "Introduction to Data Science using R"
author: "Matthew Hirschey, Ph.D."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
learnr::tutorial:
    df_print: tibble
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("inst/tutorials/plusDS_session2/code", "metadata.R")) #loads dataset-specific variables
source(here::here("inst/tutorials/plusDS_session2/code", "bespoke.R")) #loads custom objects
```

# Welcome back

# "The simple graph has brought more information to the data analyst’s mind than any other device."  -- John Tukey

## Visualize Data with
```{r ggplot_logo, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", out.height = '100%'}
url <- "https://github.com/tidyverse/ggplot2/raw/master/man/figures/logo.png"
knitr::include_graphics(url)
```
<font size = "1">[https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org)</font>

<!-- ## A basic plot using `ggplot` in R -->
<!-- We ended the last class with: -->

<!-- ```{r ggplot_basic, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE} -->
<!-- ggplot(heart, aes(x = sex, y = chol)) + -->
<!--   geom_point(color = "darkblue", size = 3) + -->
<!--   labs(x = "sex",  -->
<!--        y = "Cholesterol",  -->
<!--        title = "Cholesterol values from the heart disease dataset",  -->
<!--        caption = "Data from Kaggle | Plot from @matthewhirschey") + -->
<!--   theme_minimal() + -->
<!--   NULL -->
<!-- ``` -->

<!-- ## which built this... -->
<!-- ```{r ggplot_basic2, message=FALSE, warning=FALSE, echo=FALSE} -->
<!-- #need to parameterize this; embed rmd here?; but still need to parameterize that doc. -->
<!-- ggplot(heart, aes(x = sex, y = chol)) + -->
<!--   geom_point(color = "darkblue", size = 3) + -->
<!--   labs(x = "sex", y = "Cholesterol", title = "Cholesterol values from the heart disease dataset", caption = "Data from Kaggle | Plot from @matthewhirschey") + -->
<!--   theme_minimal() + -->
<!--   NULL -->
<!-- ``` -->

# Basic ggplot2
Any questions from last week?

## Basics of a ggplot code {.build}

Below is an example of the most basic form of the ggplot code  
  
`ggplot(data = dataframe) +`   
  `geom(mapping = aes(x, y))`  
  
Take a moment to look back at the code template.  You can see that in that code we assigned a `dataset` and the information we needed to `map` it to a type of plot

## Basics of a ggplot code {.build}
  
`ggplot(data = `r dataframe_name`) +`  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  
  
```{r simple_plot, echo = FALSE, warning = FALSE}
df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 
```

## Build a `ggplot` {.build}
- "Initialize" a plot with `ggplot()`  
  
`ggplot(data = `r dataframe_name`) +`  
  
- Add layers with `geom_functions`  
  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  

## Notes on `ggplot` style {.build}
- Generally, different people have **strong** opinions about style and data visualization  
  
- Data visualization is a rich and complex area of study and is beyond the scope of this introductory course  
  
**That being said, here are a few style tips:**    
- While you can put the `+` at the beginning of the next line, it is generally put at the end of the previous line    
  
- Arguments do not need to be explicit: this chunk of code is equivalent to the previous chunk (w/ data, mappings)  
`ggplot(`r dataframe_name`) +`    
`geom_point(aes(x = `r df_numeric2_name`, y = `r df_numeric2_name`))`  

## How are these two plots similar?
```{r double_plot, echo = FALSE, message=FALSE, warning=FALSE}
plot1 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 
plot2 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_smooth() 
plot1 + plot2
```

## The `geom` is different between these plots
`geom` is short for geometric object, which is the visual object used to represent the data  
    
`plot1 <- ggplot(`r dataframe_name`) +`  
  `geom_point(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   
  
`plot2 <-  ggplot(`r df_numeric1_name`, `r df_numeric2_name`) +`  
  `geom_smooth(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   

## Choosing your (geom)itries
Different data types require different plot types.  
  
When plotting your data, it is often helpful to take a glimpse at the data you intend to plot to know what kinds of variables you will be working with

`glimpse(`r dataframe_name`)`
```{r glimpse_again, exercise=TRUE}
```

## Geoms 
So now that you know your variable types, how do you know what geoms to use??  
  
Use the following resources to match your data type to the appropriate `geoms`

----
<a href="https://rstudio.com/resources/cheatsheets" target="_blank">https://rstudio.com/resources/cheatsheets/</a><br>
<center><iframe src="https://rstudio.com/resources/cheatsheets"></iframe></center>
----
<a href="https://www.data-to-viz.com/" target="_blank">https://www.data-to-viz.com/</a><br>
<center><iframe src="https://www.data-to-viz.com/"></iframe></center>
## Change this scatterplot code to draw boxplots {.smaller} 
`ggplot(`r dataframe_name`) + geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  
```{r boxplot_activity, exercise = TRUE}
```
  
Use the cheatsheet. Try your best guess.  
## What will this code do?
  
`ggplot(`r dataframe_name`) +`   
`geom_boxplot(aes(x=`r df_char1_name`,y=`r df_numeric1_name`)) +`  
`geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  
## Try including two `geoms` for yourself
```{r 2geom_activity, exercise = TRUE}
```
  
**Each new geom adds a new layer**  
## Aesthetics
Everything up to this point gets you a basic graph; but what about colors, shapes and overall style?  
  
You can change 5 basic aesthetics  
1. **Color**- changes the outline color of your datapoints  
2. **Size** - choose the size of the datapoint  
3. **Shape** - choose a pre-defined shape  
4. **Alpha**- changes the transparency of each point  
5. **Fill**- changes the fill color of your points  
  
Go to code/  
Open 04_ggplot2.Rmd  
Complete the exercise.    
  
## Mapping these aesthetics to data
Beyond simply changing the size or color of the variables in your plot, you can encode more information by mapping these values to data in your data set.  
  
Go to code/  
Open 05_aes.Rmd  
Complete the exercise.  
  
## Global vs Local {.build}
In ggplot2, we have the options to set mappings globally or locally. Setting a mapping globally means to set those values in the original ggplot function.  
  
Example: Earlier in class you made this graph:
`ggplot(`r dataframe_name`) +`  
  `geom_jitter(aes(x=`r df_char1_name`, y=`r df_numeric1_name`))+`  
  `geom_boxplot(aes(x=`r df_char1_name`, y=`r df_numeric1_name`))`  
  
## Global vs Local
```{r global_v_local_concepts, echo = FALSE, warning = FALSE}
ggplot(df_input) +
  geom_jitter(aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_boxplot(aes(!!sym(df_char1_name), !!sym(df_numeric1_name)))
```
## Global vs Local
However, if we map our x and y values in the ggplot function we find that we generate the same graph  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot()`  
```{r global_v_local, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot()
```
## Global vs Local {.build}
This is because when you set the aes mappings in the original `ggplot` function you are setting the `aes` **globally**.  
This means all the functions afterwards will inherit that mapping. So in our example, this means that both the jitter and boxplot geoms know to graph the same information  
  
You can also set aes values **locally** within the geom function. Doing so will only change the values in that geom
## Global vs Local
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`))`      
```{r global_v_local1, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name)))
```
## Global vs Local {.build}
```{r global_v_local2, echo=FALSE}
mean <- mean(df_numeric2_vec)
sd <- sd(df_numeric2_vec)
```
Data can also be set locally or globally. For this example, let's filter our original data first using the `dplyr::filter` function
  
`df_filter <- `r dataframe_name` %>% filter(`r df_numeric2_name` > `r round(mean + 2*sd)`)`  
  
*this number is two standard deviations above the `mean` value of `r df_numeric2_name`
## Global vs Local {.build}
Now, let's identify only the `r dataframe_about` in our data that are outliers, more than 2SD above the mean, by setting data locally in a new geom  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`   
`geom_label(data=df_filter, aes(label=`r df_id_name`))`  
## Global vs Local {.build}
```{r global_v_local3, echo = FALSE, warning = FALSE}
df_filter <- df_input %>% filter(!!sym(df_numeric2_name) > round(mean + 2*sd))
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  geom_label_repel(data = df_filter, aes(label = !!sym(df_id_name)))
```
## Global vs Local {.build}
You notice we have to indicate the new dataset, but because it has the same x and y values, we did not need to set those mappings  
  
Go to code/  
Open 06_global_v_local.Rmd  
Complete the exercise to practice mapping locally and globally.   
## Labels and Legends
Several options exist to change the default labels and legends. Recall, this code:  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`))`   
## Labels and Legends {.build}
```{r labs, echo = FALSE, warning = FALSE}
original_plot <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name)))
original_plot
```
  
But it has two problems:  
1. The x-axis label is redundant  
2. The figure legend is also redundant  
## Change labels using `labs`
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(x ="") #blank quotes removes the label`
## Change labels using `labs`
Gave us this plot:  
```{r labs_fix, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(x = "")
```
## Change legend using `guides`
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(x ="") #blank quotes removes the label +`  
`guides(color = "none")`
## Change legend using `guides` 
```{r guides, echo = FALSE, warning = FALSE}
lab_plot <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(x = "") + 
  guides(color = "none")
original_plot + lab_plot
```
## Faceting {.build}
Faceting allows you to create multiple graphs side by side in one panel. Especially useful when you want to see the data together, but not on top of each other  
  
For example:  
  
`ggplot(`r dataframe_name`) +`  
`geom_point(aes(x=`r df_char1_name`, y=`r df_numeric1_name`)) +`  
`facet_grid(cols = vars(`r df_char2_name`))`
## Faceting
```{r facetting, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_point() +
  facet_grid(cols = vars(!!sym(df_char2_name)))
```
  
*This is especially useful for exploratory data analysis
## Themes {.build}
You can change almost everything you see on your chart, but a lot of the things you may look to change are part of the "theme"  
  
Here we are going to change some features about our title text:  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12))` 
## Themes {.build}
```{r themes, echo = FALSE, warning = FALSE}
theme1 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12)) 
original_plot + theme1
```
## Themes {.build}
Next, let's change the aesthetics of our legend box  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"))`   
## Themes {.build}
```{r themes2, echo = FALSE, warning = FALSE}
theme2 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black")
  ) 
theme1 + theme2
```
## Themes {.build}
Finally, let's change the legend position  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"),`  
`legend.position = "bottom"))` 
## Themes {.build}
```{r themes3, echo = FALSE, warning = FALSE}
theme3 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black"),
        legend.position = "bottom"
  ) 
theme2 + theme3
```
## Themes {.build}
Pre-set themes also exist as an easy way to change the entire theme of your graph quickly. They can also be combined with custom theme settings  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme_minimal()`  
## Themes {.build}
```{r themes4, echo = FALSE, warning = FALSE}
theme4 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme_minimal()
theme1 + theme4
```
## Saving plots using `ggsave` {.build}
If you make a plot there are a few ways to save it, though the simplest is to use `ggsave`  
`ggsave("ggsaveexample.png", plot = last_plot())`  
You can change the type of file you save or the size.  
  
`ggsave("ggsaveexample.pdf", plot = my_awesome_object, width = 6, height = 6, units = "cm")`  
## Saving plots using ggsave
Where does it save??  
  
`getwd()`
## Putting it all together
Go to code/  
Open 07_ggplot_together.Rmd  
Complete the exercise to put all these ggplot skills to work.  
# R <br> Importing
Any questions?
## Importing data into R {.build}
The `readr` package (found in the `tidyverse` collection) contains a number of useful functions of the form `read_*` to import data. For example, if you have a `.csv` file, you would use the `read_csv` function  
  
The dataset provided to you is a cleaned R-specific document. But you will never find this in 'the wild'.  
  
Most often, you will need to find a data file (such as csv), and import it
## Importing data into R {.build}
For the purpose of this class, we have generated a **simulated** dataset of `r dataframe_join_about` to accompany the `r dataframe_name` dataset.  
To import the `r dataframe_join_file_name` file into RStudio, run the following:  
  
``r dataframe_join_name` <- read_csv(here::here("data", "`r dataframe_join_file_name`"))`

Go to code/  
Open 08_import_and_join.Rmd  
Complete the exercise to import this new dataset.  

## Importing data into R {.build}

You can also use the `readr` package to import data from a URL

For example, to load a dataset from a URL, run the following

```{r import_url, echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE}
path <- here::here("data", "phx.csv")
url <- "https://raw.githubusercontent.com/matthewhirschey/tidybiology-plusds/master/data/phx.csv"
patient_hx <- read_csv(url)
```

## Combining datasets {.build}

There are many times when you have two or more overlapping datasets that you would like to combine

The `dplyr` package has a number of `*_join` functions for this purpose

```{r left_join, echo=FALSE, fig.cap="", out.height='50%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```


## `left_join` {.build}

Returns all rows from a, and all columns from a and b

Rows in a with no match in b will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned

## `left_join` example {.build}

Take a look at the variables in each dataset - `r dataframe_name` and `r dataframe_join_name`  

You will notice that both datasets contain common variable - `r df_id_name`. This can therefore serve as a common variable to join on. Let's join on this:  
  
`left_join` `r dataframe_name` with `r dataframe_join_name` and assign the output to a new object called ``r dataframe_name`_join_left`
Go to code/  
Open 08_import_and_join.Rmd  
Complete the exercise to join the two datasets.  
  
Now you have one dataset with additional useful information
## `right_join`
```{r right_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```


## `right_join` {.build}
Returns all rows from b, and all columns from a and b

Rows in b with no match in a will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned
  
This is conceptually equivalent to a `left_join`, but can be useful when stringing together multiple steps using %>% 

## `inner_join`
```{r inner_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```

## `inner_join` {.build}
Returns all rows from a where there are matching values in b, and all columns from a and b

If there are multiple matches between a and b, all combination of the matches are returned

## `full_join`
```{r full_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```

## `full_join` {.build}
Returns all rows and all columns from both a and b  

Where there are no matching values, returns NA for the one missing    

## Scientific Reproducibly and Provenance with
```{r rmarkdown_logo, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/rstudio/hex-stickers/raw/master/PNG/rmarkdown.png"
knitr::include_graphics(url)
```
<font size = "1">[www.rstudio.com](www.rstudio.com)</font>

## R Markdown
Plain text file with 3 types of content  
```{r rmarkdown_3parts, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown.png"
knitr::include_graphics(url)
```

## Use the 'knit' button to render a report
```{r rmarkdown_knit_button, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/96ec0c54c6d64ea2ec3665db9b3b781962ff6339/5cee1/lesson-images/how-3-output.png"
knitr::include_graphics(url)
```

## How it works
```{r rmarkdown_knit, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/61d189fd9cdf955058415d3e1b28dd60e1bd7c9b/b739c/lesson-images/rmarkdownflow.png"
knitr::include_graphics(url)
```
  
When you run render, R Markdown feeds the .Rmd file to knitr ⧉, which executes all of the code chunks and creates a new markdown (.md) document which includes the code and it’s output.  
  
The markdown file generated by `knitr` is then processed by `pandoc` which is responsible for creating the finished format.  
  
This may sound complicated, but R Markdown makes it extremely simple by encapsulating all of the above processing into a single render function.  
  
## Key `knitr` points {.build}
- `Knitr` runs the document in a fresh R session, which means you need to load the libraries that the document uses **in the document**  
  
- Objects made in one code chunk will be available to code in later code chunks, but not before  
  
- For example, first create `r dataframe_name` and then using `dplyr::left_join` you create `r dataframe_joined_name`, `r dataframe_name` will be available later on in the document to do this. **However, you cannot use `r dataframe_joined_name` in a code chunk before you make it, even if it is available in your environmnet**  
  
- To keep this straight, just think (and code) in sequential chunks

## Rmarkdown -- The definitive Guide
[https://bookdown.org/yihui/rmarkdown/](https://bookdown.org/yihui/rmarkdown/)  
<center><iframe src="https://bookdown.org/yihui/rmarkdown/"></iframe></center>

# Rmarkdown <br>Text
Any questions?  

## Markdown is a simplifed language to format text
```{r rmarkdown_text, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://d33wubrfki0l68.cloudfront.net/59f29676ef5e4d74685e14f801bbc10c2dbd3cef/c0688/lesson-images/markdown-1-markup.png"
knitr::include_graphics(url)
```

## Markdown can 'markup' text to be {.build}
- **bold**  
- _italics_  
- `code`  
- unformatted text  
- ### Medium headers (3)  
- ###### Small headers (6)  
- Bullets (like this list)  
- [Links, such as to the github repository storing class material](https://github.com/matthewhirschey/)  
- ...and many more
  
## Rmarkdown cheatsheet & quick reference
<center><iframe src="https://rmarkdown.rstudio.com/lesson-15.html"></iframe></center>
  
# Rmarkdown <br>Code
Any questions?  

## Rmarkdown Code
```{r rmarkdown_code, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Shortcut
```{r rmarkdown_code_shortcut, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code-shortcut.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk1, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk1.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk2, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk2.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk3, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk3.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk4, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk4.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk5, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk5.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk6, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk6.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk7, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk7.png"
knitr::include_graphics(url)
```

## Clean-up the Rmarkdown exercise
1. Go to code/  
2. Open 10_rmarkdown.Rmd  
3. Inspect the code starting with the set-up chunk. Below it, you’ll see code that imports data, stores objects, and creates a graph.  
4. Knit the document and see the results; pretty messy, huh?  
5. Use what you've learn to execute the code, but not include the results of the set-up chunk and data import chunk.  
6. Replace every **BOLD** with inline R code  
7. Remove the {calculations} chunk so it is not included with the output  
8. Change the {graph} chunk so that only the output of the plot is shown, but not the code.
9. Re-Knit the document  
10. **Beautiful!**  

# Rmarkdown <br>YAML
Any questions?  

## Rmarkdown header is called 'YAML'
```{r rmarkdown_yaml, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-yaml.png"
knitr::include_graphics(url)
```

## Rmarkdown output is defined here
```{r rmarkdown_output, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-output.png"
knitr::include_graphics(url)
```

## Rmarkdown formats
Recall that Rmarkdown documents can be rendered into several different output file types
<center><iframe src="https://rmarkdown.rstudio.com/formats.html"></iframe></center>

## Rmarkdown Parameters
Parameters of a document are defined in the YAML header, and can pre-populate an Rmarkdown document. To see this in action,  
  
1. Open 11_rmarkdown_params.Rmd.  
2. Click the dropdown menu next to Knit and use Knit with Parameters to render the document.  
3. What happens if you type in a different gender or different age?  

--- 

```{r rmarkdown_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-parameters.png"
knitr::include_graphics(url)
```

--- 

```{r rmarkdown_using_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-using-parameters.png"
knitr::include_graphics(url)
```

## Take another look
<h1>Re-inspect 11_rmarkdown_params.Rmd.</h1>
<br>
<br>
Where were the parameters called in the code?

# Concluding Thoughts

## Data science enables {.build}
- Data science enables data-driven information gathering and hypothesis generation  
-- Scientific Research  
-- Reviews  
  
- Data science enables the ability to ask new types of questions  
  
- Process-centric, not necessarily question-centric  
  
- Making things computable makes them actionable at zero marginal cost.  
  
- Workflows save time, achieve reproducibility  

## Cheatsheet Resources
<a href="https://rstudio.com/resources/cheatsheets" target="_blank">Cheatsheets</a><br>
<center><iframe src="https://rstudio.com/resources/cheatsheets"></iframe></center>
## Stackoverflow Resources
<a href="https://stackoverflow.com/" target="_blank">Stackoverflow.com</a><br>
<center><iframe src="https://stackoverflow.com"></iframe></center>
## R4DS Resources
<a href="https://r4ds.had.co.nz" target="_blank">R for Data Science</a><br>
<center><iframe src="https://r4ds.had.co.nz"></iframe></center>
## Tidytuesday Resources
<a href="https://github.com/rfordatascience/tidytuesday" target="_blank">Tidytuesday</a><br>
<center><iframe src="https://github.com/rfordatascience/tidytuesday"></iframe></center>
## Visualization Resources: Tidytuesday
<a href="https://github.com/rfordatascience/tidytuesday" target="_blank">Tidytuesday</a><br>
<center><iframe src="https://github.com/rfordatascience/tidytuesday"></iframe></center>
## Acknowledgements {.build}
**Teaching Assistants**  
- Allie Mills, Ph.D.  
- Akshay Bareja, D.Phil.  
  
**Inspiration, ideas, packages, code**  
- R4DS (Garrett Grolemund and Hadley Wickham)  
- Mine Çetinkaya-Rundel (datasciencebox.org)  
- Chester Ismay and Albert Y. Kim (Modern Dive)  
- Garrett Grolemund  (Remastering the Tidyverse)  
- Tidyverse devs and community  
- Rstudio  
# Thank you
Any questions?
