---
title: ''
subtitle: ''
author: Matthew Hirschey, Ph.D.
date: '`r format(Sys.time(), ''%d %B, %Y'')`'
output: ~
learnr::tutorial:
  df_print: tibble
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("inst/tutorials/plusDS_session1/code", "metadata.R")) #loads dataset-specific variables
source(here::here("inst/tutorials/plusDS_session1/code", "bespoke.R")) #loads custom objects
```
# R <br>Integrated Development Environment
Any questions?

## R
Statistical Programming Language
```{r r_exercise, exercise=TRUE}

```
  
```{r r_image, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/Vb8pR6cKCRI/300×200"
knitr::include_graphics(url)
```
<font size = "1">Photo by [Photo by Evan Hein on Unsplash](https://unsplash.com/@evanheinphoto)</font>

## Rstudio
Integrated Development Environment

```{r r_de, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/XuDPnpox8tc/300×200"
knitr::include_graphics(url)
```
<font size = "1">[Photo by Jaromír Kavan on Unsplash](https://unsplash.com/@jerrykavan)</font>

---
[Open Rstudio](http://www.rstudio.cloud)  
<iframe src="http://www.rstudio.cloud"></iframe>

## Rstudio Demonstration
Go to code/  
Open 01_r_demo.Rmd  
Follow along.  
# R <br>Markdown
Any questions?

## R Markdown Introduction
Go to code/  
Open 02_rmd_exercise.Rmd  
Read through the file and do everything it tells you to do.  

## R Markdown
An authoring format for Data Science.  
```{r markdown, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/markdown.png"
knitr::include_graphics(url)
```
## Introducing `dplyr` {.build}
  
dplyr is a package that contains a suite of functions that allow you to easily manipulate a dataset  
  
Some of the things you can do are -  
  
- select rows and columns that match specific criteria  
  
- create new variables (columns)  
  
- obtain summary statistics on individual groups within your datasets  
  
The main verbs we will cover are `select()`, `filter()`, `arrange()`, `mutate()`, and `summarise()`. These all combine naturally with `group_by()` which allows you to perform any operation "by group"
